Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-10 paper/disc20clustering.tex
PPL to appendix
Notation consistency

We say that a component is \emph{trivial}, if it contains exactly one node,
and the node in this component is \emph{isolated}.
\mahmoud{we also use "singletone component". One is redundant.}
We say that a~partition is \emph{component-respecting}
if the nodes belonging to the same component are always collocated.

@@

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
$P_F := \{F_1, \dots, F_{\ell}\}$.
The \emph{distance} of a partition
 $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

\label{alg:PPL}
\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$.}
		\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Unite $C_1$ and $C_2$ into one component $C'$ and
		$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
%		\COMMENT{i.e.~if not in the same cluster.}}       
		\IF{$C_1$ and $C_2$ are not collcoated}
		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
%		\COMMENT{move to a partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}

@@

\noindent
\textbf{Component-based algorithm.}
Now we describe the algorithm \TAlg.
The algorithm partitions nodes into components, and
initially, each node belongs to its own singleton component.
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.

@@

While are bounds are tight, there are several interesting avenues for future research.
In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.

\bibliographystyle{plainurl}
\bibliography{references}
\appendix

>>>>

We say that a component is \emph{trivial}, if it contains exactly one node,
and the node in this component is \emph{isolated}.
We say that a~partition is \emph{component-respecting}
if the nodes belonging to the same component are always collocated.

@@

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
$P_F := \{F_1, \dots, F_{\ell}\}$.
The \emph{distance} of a partition
 $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).

@@

\noindent
\textbf{Component-based algorithm.}
Now we describe the algorithm \TAlg.
The algorithm partitions nodes into components, and
initially, each node is isolated (belongs to its own component).
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to trivial components, resets all counters to $0$, and ends the phase.

@@

While are bounds are tight, there are several interesting avenues for future research.
In particular, we have so far focused on deterministic algorithms, and it would be interesting to study the power of randomization in this context.
On the practical side, it would also be interesting to study our algorithms empirically, under realistic workloads.h

\bibliographystyle{plainurl}
\bibliography{references}

\appendix

\section{Pseudocode of Perfect Partition Learner}

\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a component $C_v = \{ v \}$ and add it to $\mathcal{C}$.}
		%\STATE{$P := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE{Let $C_1 \ni u$ and $C_2 \ni v$ be the  components containing $u$ and $v$, respectively.}
		\IF{$C_1 \neq C_2$}
		\STATE {Merge $C_1$ and $C_2$ into one component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$.} \label{line:mergecomponents}
		%		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
		%		\COMMENT{i.e.~if not in the same cluster.}}       
		\IF{$C_1$ and $C_2$ are not collocated}
		\STATE {Move to a partition closest to $P_I$, respecting all components in $\mathcal{C}$.} 
		%		\COMMENT{move to a partition closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}
